Missing The Point

### 2020-10-19 Manuscript.Rmd
Add description case study plots

---
title: "Missing The Point"
author: "Hanne Oberman"
date: "7-9-2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
bibliography: ref.bib
---

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
load("Data/results_25_75.Rdata")

# sim results plotting function
simulation_plot <- function(v, name, ...){
  results %>%
    mutate(
      p = factor(p, levels = c(0.25, 0.5, 0.75), labels = c("25%", "50%", "75%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR")),
      est = est - 1,
      rsq = rsq - 0.5
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}
```

@@

@@

## Example data

```{r}
# load data
load("Data/example.Rdata")
load("Data/example_mids.Rdata")

# plot each var of example data
plotfun <- function(v, ...){
ggplot(example, aes(x = .data[["it"]], y = .data[[v]])) +
  geom_line() +
  theme_classic()
}

plotfun(v="est")
plotfun(v="ac.max.qhat") #+ list(scale_y_continuous(limits = c(0.65,1)))
plotfun(v="ac.max.lambda")
plotfun(v="r.hat.max.sigma.hgt")+list(geom_hline(yintercept = 1.2, color = "grey"), scale_y_continuous(limits = c(1,2)))
plotfun(v="r.hat.max.qhat")+list(geom_hline(yintercept = 1.2, color = "grey"), scale_y_continuous(limits = c(1,2)))
plotfun(v="r.hat.max.lambda")+list(geom_hline(yintercept = 1.2, color = "grey"), scale_y_continuous(limits = c(1,2)))

plot(mids)
```

>>>>

---
title: "Missing The Point"
author: "Hanne Oberman"
date: "7-9-2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
bibliography: ref.bib
---

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
load("Data/results_25_75.Rdata")
# load data
load("Data/example.Rdata")
load("Data/example_mids.Rdata")

# sim results plotting function
simulation_plot <- function(v, name, ...){
  results %>%
    mutate(
      p = factor(p, levels = c(0.25, 0.5, 0.75), labels = c("25%", "50%", "75%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR")),
      est = est - 1,
      rsq = rsq - 0.5
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}

# plot each var of example data
casestudy_plot <- function(v, name, ...){
ggplot(example, aes(x = .data[["it"]], y = .data[[v]])) +
  geom_line(na.rm = TRUE) +
    labs(x = "Iteration number",
         y = name) +
  theme_classic()
}

@@

With this study, we show that iterative imputation algorithms can yield correct outcomes, even when a converged state has not yet formally been reached. Any further iterations would then burn computational resources without improving the statistical inferences.

Our study found that---in the cases considered---inferential validity was achieved after five to ten iterations, much earlier than indicated by the non-convergence identifiers. Of course, it never hurts to iterate longer, but such calculations hardly bring added value.

@@

## Case Study

- We use real data: the `boys` dataset from the `mice` package

- We are interested in predicting age from the other variables, in particular in the regression coefficient of `hgt`

- We compare non-convergence identified using visual inspection versus rhat in the chain variances, scientific estimate and lambda

```{r echo=FALSE}
# autocorrelation plots (not used)
# casestudy_plot(v="ac.max.qhat") #+ list(scale_y_continuous(limits = c(0.65,1)))
# casestudy_plot(v="ac.max.lambda")
```

```{r echo=FALSE}
p <- plot(mids)
p[1:2,1] # %>% plotly::ggplotly()
```

```{r echo=FALSE}
# p <- casestudy_plot(v="est", name = "Regression estimate")
# p # %>% plotly::ggplotly()
```

```{r echo=FALSE}
p <- casestudy_plot(v="r.hat.max.sigma.hgt", name = "Rhat chain variance")+list(geom_hline(yintercept = 1.2, color = "grey"), scale_y_continuous(limits = c(1,2)))
p # %>% plotly::ggplotly()
```

```{r echo=FALSE}
p <- casestudy_plot(v="r.hat.max.qhat", name = "Rhat regression estimate")+list(geom_hline(yintercept = 1.2, color = "grey"), scale_y_continuous(limits = c(1,2)))
p # %>% plotly::ggplotly()
```

```{r echo=FALSE}
p <- casestudy_plot(v="r.hat.max.lambda", name = "Rhat lambda")+list(geom_hline(yintercept = 1.2, color = "grey"), scale_y_continuous(limits = c(1,2)))
p # %>% plotly::ggplotly()
```
